---
title: "Analysis of Variance"
description: "Analyse the relationship between mutliple categorical predictor variables on a continuous response variable"
weight: 16
date: '2018-11-10'
output: blogdown::html_page
editor_options:
  chunk_output_type: console
---
```{r knitr options, message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```
```{r message=FALSE, warning=FALSE, include=FALSE}
weeds<-read.csv("../../../resources/datafiles/weeds.csv")
library(tidyverse)
weeds <- rename(weeds, flowers = flowers.m3)
frogsp<-read.csv("../../../resources/datafiles/frogs.csv", header=TRUE)
enviro<-read.csv("../../../resources/datafiles/frog_environmental.csv", header=TRUE)
insect <- read.csv("../../../resources/datafiles/insecticide.csv", header=TRUE)
```

The first step for conducting an ANOVA in R is to create an ANOVA object. There are two ways of doing this, using the `lm()` command, and using the `aov()` command. For simplicity we will be using the `aov()` command now, but we will get to the `lm()` object later.  

By using the `aov()` command, we can create an object that tells `summary()`, `plot()` or any other commands that the object is specifically for an ANOVA and as such, will be treated as one.  

The syntax for almost all analyses in R is the same. Within our analysis command (`aov()` in this example) we write a line equation for our analysis:  

  **y ~ x**  
      
This is simply your response variable (Y) and explanatory variable(s) (X) separated by a tilde **~**. The tilde acts as an = sign for the analysis. 


{{% notice note %}}
For this analysis, we will be using the **weeds** dataset.
{{% /notice %}}

```{r message=FALSE, warning=FALSE}
weeds.aov <- aov(flowers ~ species, data=weeds) # flowers (Y variable) ~ species (X variable), then data =  weeds to direct it to our dataframe. 
```

{{% notice tip %}}
As per usual, name your newly created object something that will remind you of what it is. I tend to name it something to remind me of the dataset as well as the statistical technique. Generally **dataset.analysis**. 
{{% /notice %}}

Now we have our `aov()` object, it is good practice to test the assumptions before we look at the results of the analysis. Having this workflow in place will hopefully prevent you from conveniently "forgetting" assumptions after seeing a significant result.  
