---
title: "Manipulating data part 2"
weight: 10
date: '2018-10-26'
output: blogdown::html_page
editor_options:
  chunk_output_type: console
---
```{r knitr options, message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```
```{r eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
weeds<-read.csv("../../resources/datafiles/weeds.csv")
library(tidyverse)
weeds <- rename(weeds, flowers = flowers.m3)
frogsp<-read.csv("../../resources/datafiles/frogs.csv", header=TRUE)
enviro<-read.csv("../../resources/datafiles/frog_environmental.csv", header=TRUE)
```  

#### The Select Function

This is used to select specific columns within your data and save them as a new data frame. You can use this if you have a large dataset and only want to use a few of the columns, to keep it simple and tidy. Or, you may want to take a column or two from multiple different datasets and combine them. 

```{r}
weeds_select <- select(weeds, soil) # select one column, "soil"

weeds_select <- select(weeds,c(soil, species)) # select two columns, "soil" and "species"

weeds_select <- select(weeds,c(2:4)) # select columns using numbers. In this case, select columns 2 through to 4.

weeds_select <- select(weeds, c(soil:flowers)) # select columns "soil" through to "flowers"

weeds_select <- select(weeds, -soil) # remove "soil"

# similar syntax applys for removing columns, just place a - infront e.g. select(weeds, -c(2:4))

weeds_select <- select(weeds, starts_with("s")) # select any column whose name starts with S. 
# there are many more like this above example, like "ends_with", "contains" and "matches" all which refer to the column names                    
# use the help window (?select) for more useful functions with select()

```

#### Combining data sets

One of the most frequent data manipulations for working within R is joining multiple data sets together. The most common example of this is combining species abundance (or some other variable of interest) with external sources on the environmental conditions, such as BOM data (temperature, precipitation etc.) or GPS data.  

To do most statistical analyses, data needs to be in the same data frame. So joining the datasets is an "easy" way to do so outside of excel.  

**For this exercise, we will be working with the BIOL365 Frog Data to combine the species matrix with environmental data**  

**Download the "frogs.csv" and "frog_environmental.csv"" files and read them in to R without the row.names command**  

```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
frogsp<-read.csv("frogs.csv", header=TRUE)
enviro<-read.csv("frog_environmental.csv", header=TRUE)
```

```{r}
# For a complete combine of both datasets, when there are the same number of rows in the same order
frogcombine <- bind_cols(frogsp, enviro) # In this example the "site" column has been added twice
    # There is a bind_rows() that will add rows to the bottom of a dataset, using the same syntax

frogjoin <- left_join(frogsp, enviro, by="Site") # This will join two datasets by a similar column (Site). 
# This will join the second dataset (enviro) to the first data set based on the shared column. right_join() will do the opposite

# We can use the dim() to view the dimensions of the data
dim(frogsp) # 11 columns

dim(enviro) # 16 columns
 
dim(frogjoin) # 26 colums (11 + 16 minus the 1 in common)

```

Both of these examples so far have required the same rows for each dataset. Sometimes we might have more information in one dataset then we do in the other. For this dataset we don't have this issue, so lets quickly create the issue to demonstrate.  

We will simply use the **filter()** command to filter for rows that contain a value in the "Temp" column. We have 4 rows that have an NA in "Temp" so we will use a != (not equal to) to select all rows that are not equal to NA

```{r}
enviro_filter <- filter(enviro, Temp != "NA") # This removes sites 14, 15, 35 & 36
dim(enviro_filter)

# Now we can try the two new join types

froginner <- inner_join(frogsp, enviro_filter, by="Site") # Join data. Retain only rows that occur in both data sets
dim(froginner) # 38 rows

frogfull <- full_join(frogsp, enviro_filter, by="Site") # Join data. Retain all values, all rows
dim(frogfull)

```

You can also use **semi_join()** to combine all rows that have a match in the second dataset, or **anti_join()** to combine all rows that do not match have a match in the second dataset (this ones a little weird). 


**Write Ok in the following box to continue:**  
{1:SHORTANSWER:=Ok}  