---
title: "Manipulating data part 3"
weight: 11
date: '2018-10-26'
output: blogdown::html_page
editor_options:
  chunk_output_type: console
---
```{r knitr options, message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```
```{r eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
weeds<-read.csv("../../resources/datafiles/weeds.csv")
library(tidyverse)
weeds <- rename(weeds, flowers = flowers.m3)
frogsp<-read.csv("../../resources/datafiles/frogs.csv", header=TRUE)
enviro<-read.csv("../../resources/datafiles/frog_environmental.csv", header=TRUE)
```  

#### Removing Items  

By now, we have quite a few objects in our R environment that aren't being used. The **remove()** command does exactly that....removes objects from the R environment. This helps for making things nice and tidy, specifically in our environment window. 

```{r echo=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
remove(frogfull, froginner, enviro_filter) # Removes all three objects we just generated
```

Use this in the console to "one off" remove an item.  

#### The Summarise Function  

This is an extremely useful function that lets you create different summaries of columns. You can also nest other functions within it to apply them to your columns. 

```{r}
sum_data <- summarise(weeds, mean(flowers))  # We'll start simple. Generates the mean of the flower column

sum_data <- summarise(group_by(weeds, species), mean(flowers)) 
# Using the group_by() function within summarise lets you get summaries for groups, in this case "species"

sum_data <- summarise(group_by(weeds,species, soil), mean(flowers), sd(flowers), se=sd(flowers/sqrt(n())))
# Grouped by with species & soil, generating mean, standard deviation & standard error of flowers

```

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
median <- summarise(group_by(insect, fragment), median(species.richness))
```
The last example generates the mean, sd and se for each factor combination in our dataset. This is pretty useful, particularly for generating bar graphs.  

However, its a little complex and can be in a much nicer format.  

**Use the summarise() function on the insecticide dataset to answer the following question**
**Question: In the large fragment, what is the median species richness**

{1:SHORTANSWER:=12}

#### The Pipe Function  

This lets you run multiple different functions on one dataset without having to use the intermediate steps you would have to use in base R.  
You start with the data you want to apply the functions to, followed by a pipe %>%. After each pipe you must go to the next line. 

This is useful for large messy functions with multiple nested parts. It separates everything out and makes it easier to follow.  

A pipe is simply a > nested within two percentage, %, symbols. The keyboard shortcut for this is **Ctrl** + **SHIFT** + **M**

```{r}
sum_data <- weeds %>% 
  group_by(species, soil) %>% 
  summarise(max(flowers))

# You start with the data you want to apply the functions to, followed by a pipe. After each pipe you must go to the next line   
# In this example, we grouped the data by species and soil, then performed the summarise function to generate the max number for each combination

new_data <- weeds %>% 
  mutate(binary = soil == "sandstone") %>% 
  filter(weeds == "native")

# As you can see, we can do this with most of the stuff we have already learnt
# This will generate a binary outcome (true/false) for soil with TRUE as "sandstone". Followed by filtering for "native" weeds

```

Piping is incredibly useful and much easier to read.  

**You will notice, that because we specified the data in the first line, we did not have to specify the data in the other lines, only the columns**

**Write finished in the following box to continue:**  
{1:SHORTANSWER:=finished}  
