---
title: "Columns"
weight: 9
date: '2018-10-26'
output: blogdown::html_page
editor_options:
  chunk_output_type: console
---
```{r knitr options, message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```
```{r eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
weeds<-read.csv("../../resources/datafiles/weeds.csv")
```

Another important aspect of R coding syntax is refering to specific columns. This is done by using a **$** sign after specifying our dataset and then calling the column. Like so:

```{r}
head(weeds$flowers.m3) # This says to run the head() command but only on the flowers.m3 column
```
  
Try this with some of the other commands above. *Note: Some of them will not work and will show NULL. This is because these are designed to view aspects of the data frame (e.g. names() )*  

**Now we know how to refer to a column, we can fix any issues with importing incorrect data**

```{r message=FALSE, warning=FALSE}

# Pretend for a moment our data was input incorrectly
weeds$species<-factor(weeds$species) # this would simply save the command factor() on the column species to our weeds object.

# If we wanted an ordered factor, e.g. small < medium < large we can use the following
    #example dataset
sizes <- factor(c("small", "large", "large", "small", "medium", "medium")) # creating a single column factor with 3 levels
sizes

sizes <- ordered(sizes, levels = c("small", "medium", "large")) # ordering levels from small through to large.
# Note: I did not need to specify column as this is a single column dataset. 
sizes # Now the factor is ordered.

weeds$species<-ordered(weeds$species, levels=c("Pultenaea", "Olearia", "Coprosma")) # This is an example with our data set, but this is nonsensical ordering so dont do this

```

```{r message=FALSE, warning=FALSE, include=FALSE}
weeds<-read.csv("../../resources/datafiles/weeds.csv")
```
By default, R will always sort in alphabetical order, which can be a pain when graphing. If you want ordered factors, or want to present factors along an X axis in a more logical order then the **ordered()** command or even **factor()** command where you specify levels is a good option.  

If you want to change something to a continuous (numeric, integer etc.) its a little more complicated, but in general R shouldn't mess this up too often. A quick google search or **?numeric** will help answer this. 


Once we can reference to specific columns we can do interesting things like plot a quick graph.  

```{r}
plot(weeds$flowers.m3 ~ weeds$soil) # This says to plot a graph with flowers.m3 by (~) soil. 
# You should get a box and whisker plot.  

hist(weeds$flowers.m3) # hist() is a command that only works on numeric (continuous) columns, and will show you a histogram
```
  
Try replacing the column names in the **plot()** command to see what types of graphs you get. We will return to graphing at a later date, but this is a quick and easy way to view your data. 

{{% notice question %}}
**Using the plot() command with the insecticide dataset answer the following question:**  
**Question: What fragment size has a species richness outlier? (according to the box & whisker plot)**  
{{% /notice %}}
{{% expand %}}
**Small**
{{% /expand %}}