---
title: "Two-factor Analysis of Variance"
weight: 14
date: '2018-10-26'
output: blogdown::html_page
editor_options:
  chunk_output_type: console
---
```{r knitr options, message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```
```{r message=FALSE, warning=FALSE, include=FALSE}
weeds<-read.csv("../../resources/datafiles/weeds.csv")
library(tidyverse)
weeds <- rename(weeds, flowers = flowers.m3)
frogsp<-read.csv("../../resources/datafiles/frogs.csv", header=TRUE)
enviro<-read.csv("../../resources/datafiles/frog_environmental.csv", header=TRUE)
insect <- read.csv("../../resources/datafiles/insecticide.csv", header=TRUE)
weeds.aov <- aov(flowers~species, data=weeds)

insect.aov<-aov(species.richness~fragment*spray, data=insect)
anova(insect.aov)
shapiro.test(insect.aov$residuals)
bartlett.test(species.richness~interaction(fragment, spray), data=insect)

library(car)
```
#### 4. Two-factors and Transformations

To conduct an two-factor ANOVA is pretty straightforward.
```{r}
weeds.aov2 <- aov(flowers ~ species + soil, data = weeds) # two-factor anova (without interaction)
summary(weeds.aov2)
```

This example constructs an ANOVA with two factors, but does not include the interaction term. If we want the interaction term, simply replace the **+** sign with an asterisk * .

```{r}
weeds.aov2 <- aov(flowers ~ species * soil, data = weeds) # two-factor anova (with interaction)
summary(weeds.aov2)
```

Including the asterisk tells the formula to multiply both of the factors creating the interaction factor. It will automatically produce the results for factors independantly as well as the interaction term.  

**Don't forget to check your assumptions**

Everything stays the same for assumptions except the following modifications to Bartlett's and Levene's Tests.

```{r}
bartlett.test(flowers ~ interaction(species, soil), data = weeds) # Add the interaction() argument to correctly analyse an interaction term
leveneTest(flowers ~ species * soil, data = weeds) # same syntax as the normal formula
```

**Transformations**  

There are two methods to transform your response (Y) variable for an analysis.  

1. Use a data manipulation technique such as **mutate()** to create a new column; or
2. Transform the variable within the analysis formula (see below)  

For this example, we will be log transforming the flowers column within the weeds dataset.  
**NOTE: THIS MAKES NO SENSE AS IT IS NORMAL. IT IS JUST AN EXAMPLE!**

```{r}
## Mutate Option ##
weeds <- mutate(weeds, logflowers = log(flowers)) # create new column called "logflowers"

## Formula option ##
weeds.aov.log <- aov(log(flowers) ~ species * soil, data = weeds) # log(flowers) as our Y variable tells the anova to use a log transformed response.

summary(weeds.aov.log)
```

---
**Remember**  
If you are testing assumptions, you must run the aov() (or general analysis) again with the new transformation and then extract residuals. 

```{r}
shapiro.test(log(weeds.aov$residuals)) #### DO NOT DO THIS!! ####
shapiro.test(weeds.aov.log$residuals) # Do this! #
```

See how those are different? The same thing applies to square root (sqrt) or square/cubic transformations (^2, ^3).


**Construct a Two-factor ANOVA (with interaction) on the Insecticide dataset and answer the following:**  

**1. Is the data normal?**  
{1:MULTICHOICE:Yes #Wrong~=No #OK}  

**2. What is the p-value for the Bartlett's test?**  
{1:SHORTANSWER:=0.1339}  

**3. Without transforming to normalise, what is the p-value for the interaction term?**  
{1:SHORTANSWER:=0.03275}


