---
title: "Viewing results"
weight: 20
date: '2018-11-10'
output: blogdown::html_page
editor_options:
  chunk_output_type: console
---
```{r knitr options, message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```
```{r message=FALSE, warning=FALSE, include=FALSE}
weeds<-read.csv("../../../resources/datafiles/weeds.csv")
library(tidyverse)
weeds <- rename(weeds, flowers = flowers.m3)
frogsp<-read.csv("../../../resources/datafiles/frogs.csv", header=TRUE)
enviro<-read.csv("../../../resources/datafiles/frog_environmental.csv", header=TRUE)
insect <- read.csv("../../../resources/datafiles/insecticide.csv", header=TRUE)
weeds.aov <- aov(flowers~species, data=weeds)

insect.aov<-aov(species.richness~fragment, data=insect)
summary(insect.aov)
```

Once we know our data is normal and we have our `aov()` object, we can use one of two commands on this object to generate our statistical result. The normal way to do so is to use the `anova()` command.  
```{r}
anova(weeds.aov) # run an anova on the object
# WOO! Significance!
```

Alternatively we can use the `summary()` command on our `aov()` object to generate the same result. For most other analyses, such as linear regressions and mixed models, we will use the `summary()` command exclusively.  

Because we created an `aov()` object, the `summary()` command automatically does an ANOVA.
```{r message=FALSE, warning=FALSE}
summary(weeds.aov) # print a summary of the object. In this case the summary is an anova
```

{{% notice question %}}
**Using the summary() command, what is the F-value for fragment in the Insecticide ANOVA?** 
{{% /notice %}}
{{% expand Answer %}}
14.94
{{% /expand %}}