---
title: "Facetting"
author: "Mitchell Stares"
weight: 55
date: "2018-12-03"
output: blogdown::html_page
editor_options:
  chunk_output_type: console
---
```{r knitr options, message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(tidyverse)
weeds<-read.csv("../../../resources/datafiles/weeds.csv")
weeds <- rename(weeds, flowers = flowers.m3)
weeds.summarise <- weeds %>% group_by(species) %>%
  summarise(mean = mean(flowers), se=sd(flowers)/sqrt(n()))
weeds.summarise2 <- weeds %>% group_by(species, soil) %>%
  summarise(mean = mean(flowers), se=sd(flowers/sqrt(n())))
weeds.aov2 <- aov(flowers ~ species * soil, data = weeds)
```

Facetting is the process of dividing a plot into subplots based on the values of a discrete variable. In the context of the **Iris** scatterplots, we can split by species across multiple plots.  

```{r}
ggplot(iris, aes(x=Sepal.Length, y=Petal.Length)) +
  geom_point() +
  facet_grid(.~Species)
```

As you can see, `facet_grid()` splits our single plot into a number of plots equal to the levels of the factor we specify. The syntax behind `facet_grid()` specifies our "rows" and "columns" for our plots separated by a **~** tilde. E.g. `facet_grid(row~column)`. 
If we dont have a factor for one of the column/row arguments, we just place a full stop.  

A better example of this is using `facet_wrap()` with our weeds bar graph. `facet_wrap()` sorts the graphs into a rectangular layout by default.  
```{r}
weeds.bar2 <- ggplot(weeds.summarise2, aes(x=species, y=mean, fill=species))+
  geom_bar(stat="identity", show.legend=F, colour="black")+
  labs(x="Weed Species", y= expression(Flowers~(m^3)))+
  theme(panel.background = element_blank(), panel.grid = element_blank(), axis.line = element_line(colour = "black", size=1), axis.text = element_text(colour="lightsteelblue4", size=12), axis.title = element_text(colour="steelblue", size=14, face="bold"))+
  scale_fill_manual(values = c("lightblue", "steelblue", "darkslateblue"))+
  geom_errorbar(aes(ymin = mean-se, ymax = mean+se), width=0.5)+
  ylim(0, 50)+
  facet_wrap(~soil)

weeds.bar2
```

As you can see, we have facetted by soil, creating two bar graphs of species by flowers, for each level of soil. 
With `facet_wrap()` we only specify one variable. This is useful for factors with many levels as it will automatically wrap to multiple rows. That is to say, instead of having 5 plots in a single row it might place 3 in the top row and 2 in the bottom. 