---
title: "Manipulating data"
weight: 10
date: '2018-10-26'
author: "Mitchell Stares"
output: blogdown::html_page
editor_options:
  chunk_output_type: console
---
```{r knitr options, message=FALSE, warning=FALSE, include=FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
weeds<-read.csv("../../resources/datafiles/weeds.csv")
library(tidyverse)
```  

As previously mentioned, one of the extremely useful and time saving parts of R is manipulating your data without touching your original spreadsheet. Manipulating your data within the R environment allows us to generate entirely new datasets based off our raw data, without modifying the original document. This means no more multi-sheet excel workbooks, no more opening excel to generate a new column, this can all be done in R. This is really beneficial when collaborating with other researchers, group members or supervisors since all we are required to do is send the raw data sheet and the R document. They can follow the code to see what is happening and run everything directly from the original data.  

Manipulation of data can be done with the base R language (everything we have done so far) or with packages in the **Tidyverse** library, such as **"dplyr"** and **"tidyr"**. `**Tidyverse** simplifies the language of coding and offers powerful tools for data manipulation and graphing.  

{{% notice note %}}
**Make sure you have loaded tidyverse with the `library()` command before attempting any functions.**  
{{% /notice %}}  

For the rest of the course from here on out, there will be many arguments of functions that will be left out. If you want to learn about other customisation options for your code, or are lost at any point, use the **"Help"** tab in R studio or type **"?"** followed by the name of the function.
e.g. ?rename.  

Otherwise, the internet is a awesome resource for R help.  

## The Rename Function  

By now, if you are like me, you are probably getting annoyed at writing *"flowers.m3"* or *"species.richness"* everytime you need to refer to one of those columns. This will happen ALOT with data you enter or obtain from others, as R converts any spaces to fullstops. The **rename()** function allows us to simply rename a column name within our data frame. Personally, this is my favourite function in R as I hate captials, fullstops and other annoying column name problems.  

To do this with dplyr (a tidyverse package) we simply use the following command:  

```{r}
weeds <- rename(weeds, flowers = flowers.m3)
# In the brackets we need to specify our data frame (weeds) followed by a second argument specifying the name we want for our column = the name we already have.
# Again, if you run this by itself it will not save to your data frame, unless you direct it to your data frame variable using the <- 
```

Pretty simple and straightforward.  

#### The Mutate Function  

One of the most common data manipulations is adding a new column to your dataset. This is great for transforming data, while also keeping the original. This could be used to combine multiple columns into one or perform mathematical calculations involving multiple columns with the results in a separate column.  

We will start out with a few simple methods in base R, and then move to the dplyr method.  

```{r eval=FALSE, echo=TRUE, message=FALSE, message=FALSE, warning=FALSE, include=TRUE}
##Log Transformation##

weeds$log_flowers <- log(weeds$flowers) # Base R

weeds <- mutate(weeds, log_flowers = log(flowers)) # Dplyr

# Each of these creates a new column which is the log of the flowers column.

## Basic math functions##

weeds_mutate <- mutate(weeds, flowers2 = flowers*2) 
# Simple multiplication of the flowers column by 2

weeds_mutate <- mutate(weeds, flowers_combined = flowers + flowers2) 
# This is a useless example but its just to show you how to combine multiple columns. 

weeds_mutate <- mutate(data, binary = soil == "sandstone") 
# Using boolean logic to create a column called "binary" where soil is exactly (hence double ='s) sandstone. 

weeds_mutate <- mutate(data, flowers2 = flowers*2,
                      binary = soil == "sandstone") 
# You can also perform the functions multiple times on the same data within one line. 

```

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
insect <- mutate(insect, square = species.richness^2)

```
The arguments of **mutate()** are simply the name of the data frame followed by any number of expressions that create new variables.  

You will notice throughout the **mutate()** commands that we have performed functions, creating new columns, while preserving the original. If you wish to drop the original column, simply use the **transmute()** command.

**Use the Mutate() command on the insecticide dataset to answer the following question**  
**Question: What is the square of the number in the last row of species.richness (row 42)?**  

{1:SHORTANSWER:=255}  

#### The Filter Function  

The **filter()** command is used to remove rows from your data. This can be useful for removing zeros or "no data/NA's", or for restricting certain variables in a dataset for an analysis.  

This follows the similar syntax as **mutate()** whereby we specify what dataset we want to filter, followed by how we want to filter.  

```{r eval=FALSE, echo=TRUE, message=FALSE, warning=FALSE, include=TRUE}
#The following examples will just keep overwriting the new object "weeds_filtered"

weeds_filtered <- filter(weeds, weeds == "native") 
# Gives us only the rows which are exactly "native" in the weeds column. 

weeds_filtered <- filter(weeds, weeds != "weed") 
# This gives us the same result as their are only two levels of that column. The != means "not equal to"

weeds_filtered <- filter(weeds ,flowers > 20) 
# Flowers greater than 20 m3

```

So far, we have covered renaming columns, adding new columns and filtering by rows. The next two commands are focused on selecting specific columns and creating new data tables.

**Write Ok in the following box to continue:**  
{1:SHORTANSWER:=Ok}  